---
title: "Chapters 6 to 8"
author: "Laura"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Notes for Chapter 6: Workflow: scripts 

### 6.1 Running code

Helpful keyboard shortcuts are:

* `Cmd/Ctrl + Enter` will run the complete command where your cursor is standing and take the cursor to the next command

* `Cmd/Ctrl + Shift + S` will execute the whole script


## Notes for Chapter 7: Exploratory Data Analysis

EDA is an important part of any data analysis, even if the questions are handed to you on a platter, because you always need to investigate the quality of your data.

To do data cleaning, you’ll need to deploy all the tools of EDA: visualisation, transformation, and modelling.

### 7.2 Questions

> “There are no routine statistical questions, only questionable statistical routines.” — Sir David Cox

> “Far better an approximate answer to the right question, which is often vague, than an exact answer to the wrong question, which can always be made precise.” — John Tukey

There is no rule about which questions you should ask to guide your research. However, two types of questions will always be useful for making discoveries within your data. You can loosely word these questions as:

1. What type of variation occurs within my variables?

2. What type of covariation occurs between my variables?

Tabular data is _tidy_ if each _value_ is placed in its own “cell”, each _variable_ in its own column, and each _observation_ in its own row.

```{r ch731}
diamonds %>% 
  ggplot() +
    geom_bar(aes(x = cut))
    
diamonds %>% 
  count(cut)

diamonds %>% 
  ggplot() +
    geom_histogram(aes(carat), binwidth = 0.5)

diamonds %>% 
  count(cut_width(carat, 0.5))

# if you want to go more tidyversy it can be a nightmare

diamonds$carat %>% 
  cut_width(0.5) %>% 
  as_tibble() %>%  
  count(value)

# without the warning it should be

diamonds$carat %>% 
  cut_width(0.5) %>% 
  enframe() %>%  
  count(value)

smaller <- diamonds %>% 
  filter(carat < 3)

smaller %>% 
  ggplot(aes(carat)) +
  geom_histogram(binwidth = 0.1) +
  geom_freqpoly(aes(color = cut), binwidth = 0.1)


```


## Notes for Chapter 8: