---
title: "Chapter 28"
author: "Laura"
date: "7/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes for Chapter 28: Graphics for communication

Cool extra read The truthful art by Alberto Cairo - does not seem to have an open version :_(

Come to this chapter if you need details on how to tweak your plot for nice reporting. Including details about:

* Labelling

* Annotations

* Changing and harmonizing scales

* Zooming

* All-things themes and more

* Tips for sizing your plots withing Rmarkdown and also a bit about ggsave

```{r}
df <- tibble(
  x = runif(10),
  y = runif(10),
  z = as.factor(rpois(10, 2))
)
ggplot(df, aes(x, y, size = z)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + beta + frac(delta, theta)),
    size = "Prueba"
  )
```
```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

