---
title: "Notes for Chapters 1 to 3"
author: "Laura"
date: "7/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 1: Installing the packages that will be needed

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
# install.packages(c("nycflights13", "gapminder", "Lahman"))

```

## Chapter 2: Introduction

## Chapter 3: Data visualisation

* Cool additional reading: [The Layered Grammar of Graphics](http://vita.had.co.nz/papers/layered-grammar.pdf).

* `ggplot()` creates a coordinate system to which layers are added.

* the geaom function in ggplot2 takes a ` mapping` argument. This defines how variable sin your dataset are mapped to visual properties.

```{r ggplot1, message=FALSE, warning=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```


### 3.2.4 Excercises

```{r ex3241}

ggplot(data = mpg)

mpg

?mpg

ggplot(mpg) + geom_point(mapping = aes(y = hwy, x = cyl))

ggplot(mpg) + geom_point(mapping = aes(y = class, x = drv))

```

The plot of class vs drv is not useful because all observations are one on top of another.

### 3.3 Aesthetic mappings


