---
title: "Chapters 9 and 10"
author: "Laura"
date: "11/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse); library(skimr); library(nycflights13); library(GGally); library(ggstance); library(lvplot); library(hexbin); library(modelr)

```

## Notes for Chapter 9: Intro to Wrangling 


## Notes for Chapter 10: Tibbles 

### 10.2 Creating tibbles

```{r ch1021}

as_tibble(iris)

tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)

tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb


tribble(
  ~x, ~y, ~z,
  #--|--|----
  # the previous line is there to mark where the header of the tibble is
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

### 10.3 Tibbles vs. data.frame

```{r ch1031}

tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)

```

#### 10.3.2 Subsetting

So far all the tools you’ve learned have worked with complete data frames. If you want to pull out a single variable, you need some new tools, `$` and `[[`. `[[` can extract by name or position; `$` only extracts by name but is a little less typing.

To use these in a pipe, you’ll need to use the special placeholder `.`:

```{r ch10321}

df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df %>% .$x # equivalent to df$x in the non-pipe world
df %>% .[["x"]] # equivalent to df[["x"]] in the non-pipe world

```

## 10.5 Exercises

* How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame).

* Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r ex1051}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

```

* If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

* Practice referring to non-syntactic names in the following data frame by:

1. Extracting the variable called 1.

2. Plotting a scatterplot of 1 vs 2.

3. Creating a new column called 3 which is 2 divided by 1.

4. Renaming the columns to one, two and three.

```{r ex1052}

annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

```

* What does tibble::enframe() do? When might you use it?

* What option controls how many additional column names are printed at the footer of a tibble?

## Misc

Ways to find help I am discovering:

* `vignette("<packagename>")`
* `package?<packagename>`