---
title: "Chapter 5"
author: "Laura"
date: "7/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes for Chapter 5: Data Transformation with `dplyr` 

```{r lib, include=FALSE}

library(nycflights13)
library(tidyverse)

```

A _tibble_ is a data frame slightly tweaked to work better in the tidyverse.

The six verbs of the language of data manipulation are: `filter()`, `arrange()`, `select()`, `mutate()`, `summarise()` and `group_by()`.

### 5.2.2 Logical Operators

When the logical statement gets too many parts, consider making each part a variable for code readability.

Stay away from `&&` and `||` when using `filter()`.

### 5.2.4 Excercises

```{r ex524}
flights

filter(flights, arr_delay >= 120)

filter(flights, dest == "IAH" | dest == "HOU")

airlines

filter(flights, carrier == "AA" | carrier == "DL" | carrier == "UA")

filter(flights, month <= 9 & month >= 7)

?between

filter(flights, between(month, 7, 9))

count(filter(flights, is.na(dep_time)))

filter(flights, is.na(dep_time))

NA^0
NA*0

```

### 5.3 `arrange()`

Missing values are always sorted at the end

### 5.3.1 Exercises

```{r ex531}
arrange(flights, desc(is.na(dep_time)), dep_time)

tail(arrange(flights, desc(is.na(dep_time)), dep_time))
```

### 